---
title: "LAB2"
author: "Bing Rui Yao"
date: "2022/5/2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mvtnorm)
```

# 2
```{r include=FALSE}
data<-read.table("WomenAtWork.dat",header=TRUE)
X<-as.matrix(data[,2:8])
y<-as.matrix(data[,1])
```
## a)
```{r}
mcov<-matrix(0,nrow = 7,ncol = 7)
diag(mcov)<-rep(25,7)

logpost<-function(b){
  loglike<-sum((X%*%b)*y-log(1+exp(X%*%b)))
  if (abs(loglike) == Inf){logLik = -20000}
  logPrior<-dmvnorm(b,rep(0,7),mcov,log=TRUE)
  return(loglike+logPrior)
}
b_opt<-optim(rep(0,7),fn=logpost,method="BFGS",control=list(fnscale=-1),hessian=TRUE)
b_est<-b_opt$par
J<--b_opt$hessian
```
$\tilde{\beta}$:
```{r,echo=FALSE}
b_est
```
$J_y^{-1}(\tilde{\beta})$:
```{r,echo=FALSE}
J
```

## approximate 95% equal tail posterior probability interval for the regression coefficient to the variable NSmallChild:
```{r}
c(b_est[6]-1.96*sqrt(J[6,6]/168),b_est[6]+1.96*sqrt(J[6,6]/168))
```

## Would you say that this feature is of importance for the probability that a woman works? (comparing the posterior means to the maximum likelihood estimates)
```{r}
glmModel<-glm(Work ~ 0 + ., data = data, family = binomial)
summary(glmModel)
```

## Answer: The posterior mean of coefficient of NSmallChild (-1.62277354) is close to the maximum likelihood estimates (-1.64598). The NSmallChild feature is significant for the probability that a woman works because the p-value is less than 0.001.

## b)
```{r}
x_new<-c(1,20,12,8,43,7,10)
simul<-function(n){
  beta<-rmvnorm(n,mean=b_est,sigma=J)
  p<-apply(beta,1,FUN=function(b){exp(x_new%*%b)/(1+exp(x_new%*%b))})
  return(p)
}
hist(simul(1000))
```

## c)
```{r}
x_new<-c(1,20,12,8,43,7,10)
numwork<-seq(0,11)
simul_bi<-function(n){
  beta<-rmvnorm(n,mean=b_est,sigma=J)
  p<-apply(beta,1,FUN=function(b){exp(x_new%*%b)/(1+exp(x_new%*%b))})
  p_1<-length(p[p>0.5&is.na(p)==FALSE])/n
  p_0<-length(p[p<=0.5&is.na(p)==FALSE])/n
  p_bi<-sapply(numwork,function(x){choose(11,x)*(p_1^x)*(p_0^(11-x))})
  return(p_bi)
}
plot(numwork,simul_bi(1000),type="h")
```
