---
title: "Lab 2"
author: "Hong Zhang,Bingrui Yao"
date: "18 augusti 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1
a)
Conjugate joint prior distribution:
$$ \beta|\sigma^2 \sim N(\mu_0,\sigma^2\Omega_0^{-1})$$
$$ \sigma^2 \sim Inv-\chi^2(v_0,\sigma_0^2)$$
```{r}
data<-read.table(file='TempLambohov.txt',header = TRUE)
time<-as.vector(data$time)
temp<-as.vector(data$temp)
mu0<-c(-10,100,-100)
omega0<-0.02*diag(3)
v0<-3
sigma2<-2
#simulating
ndraws<-20
beta_simul<-matrix(1,ndraws,3)
for(i in 1:ndraws){
  sigma2_simul<-rinvchisq(1,v0,sigma2)
  beta_simul[i,]<-mvrnorm(1,mu0,sigma2_simul*solve(omega0))
}
plot(x=time,y=temp,xlim = c(0,1),ylim = c(-50,50),type = 'p',col='blue',main = 'mu=(-10,100,-100)')
for(i in 1:ndraws){
  b0<-beta_simul[i,1]
  b1<-beta_simul[i,2]
  b2<-beta_simul[i,3]
  curve(b0+b1*x+b2*x**2,0,1,col='red',add=TRUE)
}
```
From the plot, the regression curves look reasonable and fit the data well, so we keep the original hyperparameters in the prior.

b)
Joint posterior diatribution:
$$ \beta|\sigma^2,y \sim N(\mu_n,\sigma^2\Omega_n^{-1})$$
$$ \sigma^2|y \sim Inv-\chi^2(v_n,\sigma_n^2)$$
```{r}
#simulating from posterior
X<-as.matrix(data.frame(cons=rep(1,length=365),x=time,x2=time**2))
beta_h<-solve(t(X)%*%X)%*%t(X)%*%temp
mu_n<-solve(t(X)%*%X+omega0)%*%(t(X)%*%X%*%beta_h+omega0%*%mu0)
omega_n<-t(X)%*%X+omega0
v_n<-v0+nrow(data)
sigma2_n<-(v0*sigma2+(t(temp)%*%temp+t(mu0)%*%omega0%*%mu0-t(mu_n)%*%omega_n%*%mu_n))/v_n

sigma_pos<-vector(length = 1000)
beta_pos<-matrix(1,1000,3)
for(i in 1:1000){
  sigma_pos[i]<-rinvchisq(1,v_n,sigma2_n)
  beta_pos[i,]<-mvrnorm(1,mu_n,sigma_pos[i]**2*solve(omega_n))
}
```
i)Plot a histogram for each marginal posterior of the parameters:
```{r}
hist(sigma_pos,main = 'histogram for sigma2',xlab = 'sigma2')
hist(beta_pos[,1],main = 'histogram for beta0',xlab = 'beta0')
hist(beta_pos[,2],main = 'histogram for beta1',xlab = 'beta1')
hist(beta_pos[,3],main = 'histogram for beta2',xlab = 'beta2')
```
ii)Make a scatter plot of the temperature data and overlay a curve for the posterior median:
```{r}
plot(time,temp,type = 'p',col='blue',ylim = c(-30,30))
lines(time,apply(X%*%t(beta_pos),1,median),col='green')
lines(time,apply(X%*%t(beta_pos),1,quantile,prob=0.025),col='red')
lines(time,apply(X%*%t(beta_pos),1,quantile,prob=0.975),col='red')
```
The posterior probability intervals doesn't contain most of the data points.It shouldn't contain most of the data points because it's the 95% equal tail probability intervals of the expected value of temperature, which doesn't contain the effect of $\sigma^2$.

c)
```{r}

```
d)
```{r}

```